---
title: Bar-chart demo
output: html_document
---

```{r}
library("ggplot2")
library("ggvega")
library("tibble")
library("htmltools")
library("purrr")
library("jsonlite")
library("vegawidget")
```

Let's start out with an initial dataset:

```{r data-init}
set.seed(314)

n_obs <- 100

lang_prob <- tribble(
     ~language, ~probability,
           "R",         0.40,
      "Python",         0.30,
         "C++",         0.15,
  "JavaScript",         0.10,
        "Lisp",         0.05
)

data_all <- 
  tibble(
    user_id = seq(1, 2 * n_obs),
    language = 
      sample(
        lang_prob$language, 
        size = 2 * n_obs, 
        replace = TRUE, 
        prob = lang_prob$probability
    )
  )  

data_init <- data_all[data_all$user_id <= n_obs, ]
data_new <- data_all[data_all$user_id > n_obs, ]
```

```{r}
g <- ggplot(data_init, aes(x = language)) + geom_bar()
g
```

We create a vegaspec based on the ggplot2 design. We use the `vegawidget()` function because we want to specify the element ID.

```{r}
vs <- as_vegaspec(g)

vegawidget(vs, elementId = "vis")
```

Create an HTML control:

```{r control}
control <- 
  tags$div(
    tags$span("dataset"),
    tags$select(
      name = "dataset",
      tags$option(value = "data_init", "Initial"),
      tags$option(value = "data_new", "New")
    )
  )

control
```

We need to get the data-frames into JavaScript somehow.

```{r data}
ser <- function(x) {
  x <- purrr::transpose(x)
  x <- jsonlite::toJSON(x, auto_unbox = TRUE)
  
  x
}

data_script <- 
  tags$script(
    type = "text/javascript",
    glue_js(
      "var data_all = {'data_init': ${ser(data_init)}, 'data_new': ${ser(data_new)}};"
    )
  )

data_script
```

Next, we define what we want to happen when the `dataset` input changes:

```{js}
// {js}
function on_data() {

  // get a local dataset
  var data_local = data_all[data_choice.value];
  
  // change Vega chart
  Vegawidget.findViewPromise('#vis').then(function(view) {
    var changeSet = vega.changeset()
                        .insert(data_local)
                        .remove(vega.truthy);
                          
    view.change('data-00', changeSet).run();  
  });  
}
```

Finally, we set a listener to implement what we want to happen when the `dataset` input changes:

```{js}
// {js}
var data_choice = document.querySelector("[name=dataset]");

// whenever the input changes, run the updating function
data_choice.addEventListener("input", on_data);

// run the updating function *once* to initialize the input to the data
on_data();
```
