---
title: bar chart
output: github_document
---

```{r message=FALSE}
library("ggvega")
library("ggplot2")
library("vegawidget")
```

The purpose of this document is to show how bar-chart examples can be translated.

## "Simple" bar chart

The is the simplest bar chart we can imagine. On the `x`-axis, we have the `class` of the car; on the `y`-axis we have the number of cars (obsewrvations) in each class.

Also note that I have specified the `width` and `height` of the *rendering* of the vegaspec; this is *not* part of the translation.

```{r results='asis'}
dev_gallery("barchart-mpg")
```

One thing to note is that the `stat` contains a `default_aes` elementm which we would use in the absence of an explicit aesthetic-mapping for `y`:

```{r}
g <- last_plot()

g$layers[[1]]$stat$default_aes
```

For example, we could define the aesthetic mapping for `y`:

```{r}
  g <- ggplot(data = mpg, aes(x = class, y = stat(count) + 100)) +
  geom_bar()

  g
```

This formulation specifies an aesthetic mapping using (what we call) a `stat_expression`, which we do not yet support.

```{r}
g$mapping$y
```

However, we would be able to support (once we support `StatBin`) specifying the type of stat explicitly, using its `default_aes`.


```{r}
g <- ggplot(data = mpg, aes(x = displ)) +
  geom_bar(stat = "bin")

g
```

In short, for now I think that we can use only the `default_aes` for the `y` aesthetic, but we need to keep in mind that it could be overriden by a `stat_expression` in the future. We will have to handle the `weight` aesthetic as a `field`.

What can we do now to support `stat(count)` in the mapping?

### Translation

For `StatCount`, we build a `y` encoding:

- add `type: "quantitative"`
- if `weight` is a value, add `agg: "count"`
- if `weight` is a field, add `agg: "sum", field: "<field>"`

For `PositionStack`, we add to the `y` encoding:

- `stack: "zero"`

For `PositionFill`, we add to the `y` encoding:

- `stack: "normalize"`

We should also note that the functions `position_stack()` and `position_fill()` each have two parameters:

- `vjust`: numeric, default `1`
- `reverse`: logical, default `FALSE`

For the time being, we will not handle these parameters. (I suspect that we can use vjust only with `geom_text()`, so I don't look forward to having to handle that interaction between `Geom` and `Position`)

It seems we have more examples to handle:

## Stacked bar-chart

```{r results='asis'}
dev_gallery("barchart-stacked-mpg")
```

## Weighted bar-chart

```{r results='asis'}
dev_gallery("barchart-weighted-mpg")
```

## Normalized bar-chart

```{r results='asis'}
dev_gallery("barchart-normalized-mpg")
```

## Flipped, stacked, weighted, normalized, bar-chart

```{r results='asis'}
dev_gallery("barchart-flipped-weighted-normalized-mpg")
```
