---
title: "Histograms & Bar Charts"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("vegawidget")
library("ggplot2")
```

### Histograms in ggplot2

`geom_histogram()` displays the counts with bars and uses `stat_bin()` by default. 

`stat_bin()` is suitable only for continuous x data. `stat_count()` can be used if the x data is discrete.

`geom_histogram()` uses the same aesthetics as `geom_bar()`

### Histograms in Vega-Lite


Binning can be done in either encoding or transform. It is also an option to use data that is already binned.

The additions to be noted are `"bin": true` to the x encoding and the `"aggregate": "count"` to the y encoding.


```
{
  "data": {},
  "mark": "bar",
  "encoding": {
    "x": {
      "bin": true,
      "field": "",
      "type": "quantitative"
    },
    "y": {
      "aggregate": "count",
      "type": "quantitative"
    }
  }
}

```

### Histogram examples


##### A simple histogram

```{r}
ggplot(diamonds, aes(carat)) +
  geom_histogram()
```

```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = diamonds),
  mark = "bar",
  encoding = list(
    x = list(bin = list(maxbins = 30), field = "carat", type = "quantitative"),
    y = list(aggregate = "count", type = "quantitative")
  )
) %>% 
  as_vegaspec()
```


## Bar charts

### Bar charts in ggplot2

There are two types of bar charts in ggplot2: `geom_bar()` and `geom_col()`.

- `geom_bar()` makes the height of the bar proportional to the number of cases in each group (or if the `weight` aesthetic is supplied, the sum of the weights).  

  - uses `stat_count()` by default: it counts the number of cases at each x position  
  

- `geom_col()` makes the heights of the bars to represent values in the data.  

  - uses `stat_identity()` by default: it leaves the data as is.

### Bar charts in Vega-Lite

The important addition to be noted is the `"aggregate": "count"` added to the y encoding.

```
{
  "data": {},
  "mark": "bar",
  "encoding": {
    "x": {
      "field": "",
      "type": "nominal" 
    },
    "y": {
      "aggregate": "count",
      "type": "quantitative"
    }
  }
}

```

If a weighting variable is used in ggplot2, then we would specify the variable name as the field in the y encoding and the aggregate function is now `sum` instead of `count`.

```
{
  "data": {},
  "mark": "bar",
  "encoding": {
    "x": {
      "field": "",
      "type": "nominal" 
    },
    "y": {
      "aggregate": "sum",
      "field": "",
      "type": "quantitative"
    }
  }
}

```

### Bar chart examples

##### A bar chart

```{r}
ggplot(mpg, aes(class)) + geom_bar()
```

```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = mpg),
  mark = "bar",
  encoding = list(
    x = list(field = "class", type = "nominal"),
    y = list(aggregate = "count", type = "quantitative")
  )
) %>% 
  as_vegaspec()
```





<br/>

##### A bar chart with a weight variable


```{r}
ggplot(mpg, aes(class)) + geom_bar(aes(weight = displ))
```

```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = mpg),
  mark = "bar",
  encoding = list(
    x = list(field = "class", type = "nominal"),
    y = list(aggregate = "sum", field = "displ", type = "quantitative")
  )
) %>% 
  as_vegaspec()
```



<br/>

##### A stacked bar chart


```{r}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv))
```

```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = mpg),
  mark = "bar",
  encoding = list(
    x = list(field = "class", type = "nominal"),
    y = list(aggregate = "count", type = "quantitative"),
    color = list(field = "drv", type = "nominal")
  )
) %>% 
  as_vegaspec()
```


<br/>

##### A flipped stacked bar chart


```{r}
ggplot(mpg, aes(class)) +
 geom_bar(aes(fill = drv)) +
 coord_flip() +
 theme(legend.position = "top")

```

```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = mpg),
  mark = "bar",
  encoding = list(
    y = list(field = "class", type = "nominal"),
    x = list(aggregate = "count", type = "quantitative"),
    color = list(field = "drv", type = "nominal")
  )
) %>% 
  as_vegaspec()
```

<br/>

##### A grouped bar chart using `position_dodge()`

```{r}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv), position = position_dodge())
```


<br/>

##### A grouped bar chart using `position_dodge2()`

```{r}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv), position = position_dodge2())
```


```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = mpg),
  mark = "bar",
  encoding = list(
    column = list(field = "class", type = "nominal"),
    x = list(field = "drv", type = "nominal",
      scale = list(rangeStep = 12),
      axis =  list(title = "")),
    y = list(aggregate = "count", type = "quantitative"),
    color = list(field = "drv", type = "nominal")
  ),
  config = list(
    view = list(stroke = "transparent"),
    axis = list(domainWidth = 1)
  ),
  spacing = 6
) %>% 
  as_vegaspec()
```


<br/>

##### A normalized bar chart (showing proportions) using `position_fill()`


```{r}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv), position = position_fill())
```

```{r}
list(
  `$schema` = vega_schema(), 
  data = list(values = mpg),
  mark = "bar",
  encoding = list(
    x = list(field = "class", type = "nominal"),
    y = list(aggregate = "count", type = "quantitative", stack = "normalize"),
    color = list(field = "drv", type = "nominal")
  )
) %>% 
  as_vegaspec()
```

<br/>

##### An ordered bar chart 


```{r}


```


<br/>



### Next steps

How to translate stat params?
