---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "svg"
)
```

# ggvega

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ggvega)](https://cran.r-project.org/package=ggvega)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/vegawidget/ggvega.svg?branch=master)](https://travis-ci.org/vegawidget/ggvega)
<!-- badges: end -->

The goal of **ggvega** is to translate a ggplot2 object to a Vega-Lite specification.

``` {r example}
library(ggvega)
library(ggplot2)

p <- 
  ggplot(iris) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length, colour = Species))

ggspec<-gg2spec(p)

ggvega(ggspec)


```

## test
```
library(ggvega)
library(ggplot2)

p <- 
  ggplot(iris) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length, colour = Species))

as_vegaspec(p)
```

### Important Note

This package is in an *extremely* experimental state, under *very* active development. 

At present, we can translate:

- `geom_point()`
- aesthetics: `x`, `y`, `colour`
- unit-plots; no facetting.

We have plans to expand this capability; presently, we are working out a system to add new capabilities quickly and reliably.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vegawidget/ggvega")
```
