---
title: Deploying examples
output: github_document
---

The motivation for this effort is to do **verification** and **validation**. The [definition provided by Barry Boehm](https://en.wikipedia.org/wiki/Verification_and_validation#Overview) is useful:

- **verification** asks the question "are we building the thing right?"

- **validation** asks the question "are building the right thing?"

Accordingly, the purpose of this document is to manange the deployment of examples to two places:

- a testing directory, where we **verify** that each of the ggplot examples creates the ggspec and vegaspec we expect. 

- a package-examples directory, containing examples of ggplots that we claim can be translated. These examples would be the source of a pkgdown gallery. This gallery will also serve as a visual regression, to **validate** that each translation appears as we expect.

A third use of these examples is to create development-documentation.

Our examples appear in the contexts of ggplot2, ggspec, and Vega-Lite. In the interests of consistency, the directory for a particular type of example is named after the type of object it creates:

```
ggplot/
  scatterplot-iris.R
  ...
ggspec/ 
  scatterplot-iris.R
  ...
vegaspec/
  scatterplot-iris.R
  ...
```

Each example consists of the R source code used to create the particular object. We, the developers, create these examples by hand, as we specify new features.

In the R code, each example is referred-to using the `example` keyword. Above, `example ` is `"scatterplot-iris"`.


```{r setup}
library("conflicted")
library("magrittr")
library("ggvega")
library("purrr")
library("here")
library("fs")
```

```{r source}
source_dir <- here("data-raw", "examples", "ggplot2")
source_files <- fs::dir_ls(source_dir, regexp = "[.]R$")
```

## Testing examples

This is where we build our testing examples, which are created in `tests/testthat/examples/`:

```
ggplot/
  scatterplot-iris.R
  ...
ggspec/ 
  scatterplot-iris.gg.json
  ...
vegaspec/
  scatterplot-iris.vl.json
  ...
```

Here, we make a few changes to the original examples:

- all the data in the examples are truncated. By this, we mean that each data frame is each example is modified so that it keeps only its first row. The size of the data-frame has no effect on the rest of the specification. By truncating the data frame, can make the test lighte, and each of the examples more-understandable to humans (less JSON to deal with).

- the ggspec and vegaspec examples are stored as JSON, rather than as the source-code used to create them.

As described in at the top of this document, when we develop new cases we will manually create new examples for ggspec and vegaspec to show how the feature is to be translated. Because we will develop the examples before we develop the capability to make the translation, we will want to skip certain examples from being tested.

As such, we have two collections:

- `skip_ggspec`: examples for which we do not yet have a `ggspec` translation
- `skip_vegaspec`: examples for which we do not yet have a `vegaspec` translation

Examples that appear in a skip collection are not copied into that testing directory.

```{r skip}
# Use this chunk to moderate which examples are copied into the testing directories.

skip_ggspec <- c()

skip_vegaspec <- c()

# if something is skipped for ggspec, it is also skipped for vegaspec
skip_vegaspec <- c(skip_ggspec, skip_vegaspec)
```


As above, we create an empty directory to contain the testing examples.

```{r clean-testing}
testex_dir <- here("tests", "testthat", "examples", "ggplot2")
testex_dir

if (dir_exists(testex_dir)) {
  dir_delete(testex_dir)  
}

dir_create(testex_dir)
```

Next, we need a function that will:

- read a file into lines
- for each line, substitute the `data` argument with `head(data, 1)`
- write those lines to a new file, in the test directory

```{r}
transfer_test <- function(path) {
  
  new_path <- fs::path_join(c(testex_dir, basename(path)))
  
  content <- readr::read_lines(path)
  content <- purrr::map_chr(content, ggvega:::head_data)

  readr::write_lines(content, path = new_path)
}
```

```{r}
walk(source_files, transfer_test)
```

## Package examples

First, we create an empty directory to contain our package examples.

```{r clean-package}
pkgex_dir <- here("inst", "examples", "ggplot")
pkgex_dir

if (dir_exists(pkgex_dir)) {
  dir_delete(pkgex_dir)  
}

dir_create(pkgex_dir)
```

Next, we copy all the files from the source directory to the package-example directory:

```{r copy-examples}
walk(source_files, fs::file_copy, new_path = pkgex_dir)
```

Finally, to confirm, we list the files in the package-source directory

```{r list-examples}
dir_ls(pkgex_dir) %>% basename()
```



## Adding a new example
