---
title: "First components of the ggspec function"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
library("ggplot2")
library("dplyr")
```

```{r utils, echo = FALSE}
"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}

"%|W|%" <- function(a, b) {
  if (!is.waive(a)) a else b
}

is.waive <- function(x) inherits(x, "waiver")
```


<br/>



## Extracting elements

<br/>

```{r}
p <- ggplot(data = iris) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length, color = Species), shape = 21, fill = "white") +
  scale_y_log10()

p
```

<br/>




### Extracting data

The first move will be to create the list `data` where all of the data will live. 

<!-- QUESTION: what to do about `waiver()` objects? -->

#### Intermediate data step

`data_int()` will create an intermediate-form for the data. The inputs are the plot data and the plot layers. The result is a named list of datasets, named `data-00`, `data-01`, ..., where each list contains the elements `metadata`, `variables`, and `hash`.

**TO - DO**:  

- Need to name the datasets but will need to check for matching hashsums and remove `NULL`s  
  - Would like only the default data to be able to be named `data-00`  
  - The other data sets will be named `data-01`, `data-02`, and so on.  
- `metadata` will be a named list (names are variable names) of 3 (1 required, 2 optional)
  - `type` = variable type (required)
  - `levels` = levels of factor (optional)
  - `timezone` = timezone of `date` or `POSIXct` (optional)

```{r}
name_data <- function(dat) {
  # check for any matching hashsums
}

data_int <- function(data_plt, layers_plt) {
  #join together default data and layer data
  data_all <- append(list(data_plt), purrr::map(layers_plt, purrr::pluck, "data"))
  
  #format the lists of data
  data_all <- purrr::map(data_all, format_data_int)
  
  # how to name the datasets??
  
  # remove NULL entries 
  data_all <- purrr::discard(data_all, is.null)
  
  
  
  
  data_all
}
```

<br/>

Helper functions:

`format_data_int()` will format each list of data so that it contains:  

- `metadata`: discussed below
- `variables`: the data frame itself  
- `hash`: the md5 hash of the data frame


```{r}
format_data_int <- function(dat) {
  if(is.waive(dat) || is.null(dat)) return(NULL) 
  else {
    list(
      metadata = purrr::pluck(dat) %>% purrr::map(create_meta_levels),
      variables = dat,
      hash = digest::digest(dat)
    )
  }
}
```

`create_meta_levels()` will create the names list of 3

- `metadata`: could be a named list, using names of variables:  
    - `type`: first pass at `"quantitative"`, ..., based on class, etc.  
    - `levels`: optional, vector of strings, used for factor-levels  
    - `timezone`: optional, timezone of `date` or `POSIXct`

`case_type_vl()` converts the type into a Vega-lite type

```{r}
case_type_vl <- function(type) {
  case_when(
    type == "Date" | type == "POSIXct" ~ "temporal",
    type == "factor" | type == "character" | type == "logical" ~ "nominal",
    type == "ordered" ~ "ordinal",
    type == "numeric" ~ "quantitative"
  )
}

create_meta_levels <- function(dat){
  type = class(dat)
  if(type == "factor" | type == "ordered") {
    meta <- list(
      type = case_type_vl(type),
      levels = levels(dat)
    )
  } else if (type == "date" | type == "POSIXct") {
    meta <- list(
      type = case_type_vl(type),
      timezone = NULL # use lubridate::tz or ??
    )
  } else {
    meta <- list(
      type = case_type_vl(type)
    )
  }
  meta
}

```

<br/>

Example of the function in use:

```{r}
test <- data_int(p$data, p$layers)
str(test)
```

<br/>

This intermediate-form could be used to generate the ggspec-form; it could also be useful later.

<br/> 

#### Final data step


`data_spc()` will return a named list of datasets, named `data-00`, `data-01`, ... . Each list will have:  

- `metadata`, as in `data_int()`  
- `observations`, transpose of variables  

```{r}
format_data_spec <- function(dat) {
  list(
    metadata = dat$metadata,
    observations = purrr::transpose(dat$variables)
  )
}

data_spc <- function(data_int) {
  purrr::map(data_int, format_data_spec)
}
```

```{r}
str(data_spc(data_int(p$data, p$layers)), max.level = 2)
```


<br/>



### Extracting layers

Within each layer-object, we need:  

1. data (a reference id?)  
2. geom  
3. geom_params (maybe)  
4. mapping  
5. aes_params  
6. stat (maybe)  
7. stat_params (maybe)  


<br/>

The ggspec layers are a function of the ggplot layers, but also of the data:

`layer_spc()` calls `get_layers()` for each layer. `get_layers()` returns ...

- if `layer_plt` has no data, use `data-00`  
- if `layer_plt` has data, hash it and compare against `data_int`, use name  
- make sure that the mapping field is a name in the dataset  


```{r}
layer_spc <- function(layer_plt, int_data) {
  purrr::map(layer_plt, get_layers, int_data)
}
```

<br/>

Helper functions:

```{r}
get_layers <- function(layer, int_data) {
  pluck_layer <- purrr::partial(purrr::pluck, .x = layer)
  
  list(
    data = pluck_layer("data") %>% compare_data(int_data),
    geom = list(
      class = pluck_layer("geom", class, 1)
    ),
    mapping = pluck_layer("mapping") %>% purrr::map(get_mappings),
    aes_params = pluck_layer("aes_params"),
    stat = list(
      class = pluck_layer("stat", class, 1)
    )
  )
}
```


```{r}
get_mappings <- function(aes) {
  list(field = rlang::get_expr(aes)) 
}
```

In `compare_data()`:

- if `layer_plt` has no data, use `data-00`  
- if `layer_plt` has data, hash it and compare against `data_int`, use name  
- make sure that the mapping field is a name in the dataset  

```{r}
# 
compare_data <- function(layer_data, plot_data){
  NULL
}
```


<br/>

Example of function being used:

```{r}
str(layer_spc(p$layers, test))
```

<br/>


### Extracting scales  


I think that scales will be one-to-one:

`scales_spc()` calls `get_scales()` which operates on a single scale, used with purrr::map(), to get ...

will need to first check if there is even anything there...

```{r}

get_scales <- function(scale) {
  
  pluck_scale <- purrr::partial(purrr::pluck, .x = scale)
    
  list(
    name = pluck_scale("name"),
    class = pluck_scale(class, 1),
    aesthetics = pluck_scale("aesthetics"),
    transform = list(
      name = pluck_scale("trans", "name")
    )
  )
}

scale_spc <- function(scale_plt) {
  purrr::map(scale_plt, get_scales)
}
```

```{r}
str(scale_spc(p$scales$scales))
```


<br/>


### Extracting labels  


Finally, labels:


```{r}
find_scale_labs <- function(labs) {
  lab <- purrr::pluck(labs, "name")
  if(!is.waive(lab)) {
    names(lab) <- purrr::pluck(labs, "aesthetics", 1)
    lab
  }
  
}

labels_spc <- function(labels_plt, scales_plt) {
  # Find the right way to deal with labels - we could run into a
  # problem if we have, say, multiple color scales
  
  # scale_labs <- purrr::map(p_scale$scales$scales, find_scale_labs)
  
  # How to replace the labels with scale labels???
  
  labels_plt
}


```



```{r}
p_lab <- ggplot(iris) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length), color = "firebrick") +
  scale_y_log10() +
  labs(x = "new lab")

p_scale <- ggplot(iris) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length), color = "firebrick") +
  scale_y_log10("new lab") 

ps <- ggplot_build(p_scale)
```


<br/>


## All together now!

<br/>

```{r}
ggspec <- function(plt){
  list(
    data = data_spc(data_int(plt$data, plt$layers)),
    layers = layer_spc(plt$layers),
    scales = scale_spc(plt$scales$scales),
    labels = labels_spc(plt$labels)
                    
  )
}
```

<br/>


Try it out:

```{r}
str(ggspec(p), max.level = 3)


```


<br/>
<br/>

```{r eval = FALSE}
pp <- ggplot(iris) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length, color = Species)) + 
  geom_point(aes(x = Petal.Width, y = Petal.Length), color = "firebrick", size = 1) +
  scale_y_log10("new lab") 

str(ggspec(p_scale), max.level = 3)
str(ggspec(pp), max.level = 3)


```





